Towards an Automated Sportscasting System for Chess

### 2011-05-18 paper/paper (copy).tex
Broken paper

\section{Introduction}
Chess has a long history of intriguing enthusiasts due to its complexity. Even after years of play, each game is new challenge, a new problem to solve. Historically, chess has been prominent as a sport where the battles of grandmasters engulf thousands; given the highly intricate nature of chess (and other similar abstract strategy games), experts have scrutinized these games to devise a standard high-level structures like openings, strategies, tactics, and time advantage, which they use to explain maneuvers and blunders that occur in the games. Devotees of chess also attempt to improve their own performance in terms of these high-level structures, reading chess theorists' comments about games by high-ranking players.

@@

\section{Related Work}
\begin{itemize}
\item Different applications of similar techniques
\item Differences from sportscasting for, e.g., RoboCup
\item Text about approach
\end{itemize}

@@

\paragraph{Automated sportscasting.}

\paragraph{Automatic chess commentary.} There exist techniques to automatically annotate chess games which are heavily dependent on rule-based logic and use of a chess-engine to determine which parts of a specific chess game are salient~\cite{cambridge-chess-annotation}. As a departure from the rule-based approach to automatic chess commentary, Guid et al. (\cite{learning-positional-features}) discuss the use of argument based machine learning, which analyzes reasoning where arguments for and against a certain claim are produced and evaluated (in the context of chess, why a certain move was played over others.) Their system, much like the one presented in this paper, identifies complex positional features of a given chess game.

However, we identified two shortcomings of the AMBL system.  First, it relies on experts to provide arguments as to why a certain position exemplifies (or doesn’t) a particular positional feature.  While this is a step up from defining a classification rule by hand, it is infeasible to iterate this procedure multiple times.  Moreover, Guid described that

@@

\subsection{Data Pre-Processing}	
In order to ensure consistency and integrity over the games obtained from disparate sources, we sanitized them in multiple ways. We parsed each game in the data set and removed information that would confuse the parser. For example, some games had English or German comments by humans which we did not plan to use, and sometimes resulted in parsing errors due to specific characters present in the comments. Others had alternate branches, sometimes up to multiple levels (alternate branches within alternate branches), which the parser could not handle properly. Still others represented games in a non-standard way (for example, pawn promotions were represented without the ``=''), which we had to fixed.

@@

\subsection{Parsing}
We tried various open source PGN parsers, each succumbing to flaws like mishandling uncommon cases (like promotions and special moves like en passant). We based our system on Chesspresso,\footnote{url{http://www.chesspresso.org/}} a Java library for writing chess programs, which we fixed and modified for our needs. Chesspresso takes a PGN file as input and returns a data structure that represents the initial position of the board and the subsequent moves (as well as all metadata necessary to recreate the PGN file). We also extended it to be more robust to variations in input format, in addition to being able to export PGNs in a canonical form, which was crucial for removing duplicates.

\section{Features}
\subsection{Manually Generated Features}
As a first step towards classification, we selected chess-specific manual features from a given position-move pair.  The position-based features involve information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move is a capturing move or a check.  We list all our hand-selected manual features in Table REF.

\begin{table*}
\begin{tabular}{lp{0.6\textwidth}}
\hline
\textbf{Category} & \textbf{Description} \\ \hline
Position features & Piece location, potential moves per piece, material count, square control. \\
Move features & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move.
\end{tabular} 
\caption{List of chess-specific manual features per position.}
\end{table*}

\subsection{Sparse Autoencoding}

@@

@@

We then extended the tagging to also incorporate information about the player

@@

\subsection{Classification Quality}

>>>>

\section{Introduction}
Chess has a long history of intriguing enthusiasts due to its complexity. Even after years of play, each game is new challenge, a new problem to solve. Historically, chess has been prominent as a sport where the battles of grandmasters engulf thousands; given the highly intricate nature of chess (and similar abstract strategy games), experts have scrutinized these games to devise a standard high-level structures like openings, strategies, tactics, and time advantage, which they use to explain maneuvers and blunders that occur in the games. Devotees of chess also attempt to improve their own performance in terms of these high-level structures, reading chess theorists' comments about games by high-ranking players.

@@

\section{Related Work}

@@

\paragraph{Automated sportscasting.} Chen and Mooney (\cite{learning-to-sportscast}) present a commentator system that learns language from sportscasts of simulated RoboCup games. Their system, which is trained on ambiguous supervision in the form of textual human commentaries and simulation states of the soccer games, simultaneously tries to establish correspondences between the commentaries and the simulation states. Their work focuses on generating meaningful English commentary from relatively simple topics such as ‘Player 1 kicked the ball to Player 2’, while our work focuses more on identifying complex topics.

\paragraph{Automatic chess commentary.} There exist techniques for automatic annotation of chess games which are heavily dependent on rule-based logic and use of a chess-engine to determine which parts of a specific chess game are salient~\cite{cambridge-chess-annotation}. As a departure from the rule-based approach to automatic chess commentary, Guid et al. (\cite{learning-positional-features}) discuss the use of argument based machine learning, which analyzes reasoning where arguments for and against a certain claim are produced and evaluated (in the context of chess, why a certain move was played over others.) Their system, much like the one presented in this paper, identifies complex positional features of a given chess game.

However, we identified some shortcomings of the AMBL system. First, it relies on experts to provide arguments as to why a certain position exemplifies (or doesn’t) a particular positional feature. While this is an improvement from defining a classification rule manually, it is infeasible to repeat this procedure multiple times. Second, the arguments that the experts are allowed to provide rely on a preset list of features.

@@

\subsection{Data Pre-Processing}	
In order to ensure consistency and integrity over the games obtained from disparate sources, we sanitized them in multiple ways. We parsed each game in the data set and removed information that would confuse the parser. For example, some games had English or German comments by humans which we did not plan to use, and sometimes resulted in parsing errors due to specific characters present in the comments. Others had alternate branches, sometimes up to multiple levels (alternate branches within alternate branches), which the parser could not handle properly. Still others represented games in a non-standard way (for example, pawn promotions were represented without the ``=''), which we had to fix.

@@

\subsection{Parsing}
We tried various open source PGN parsers, each succumbing to flaws like mishandling uncommon cases (such as  promotions and special moves like \emph{en passant}). We based our system on Chesspresso,\footnote{url{http://www.chesspresso.org/}} a Java library for writing chess programs, which we fixed and modified for our needs. Chesspresso takes a PGN file as input and returns a data structure that represents the initial position of the board and the subsequent moves (as well as all metadata necessary to recreate the PGN file). We also extended it to be more robust to variations in input format, in addition to being able to export PGNs in a canonical form, which was crucial for removing duplicates.

\subsection{Preparing the Data}
After parsing each game, we filtered out the tagged games that had more than 10-plies
JINGHAO: Write about slicing and tag generation

\section{Features}
\subsection{Sparse auto-encoder}
The sparse auto-encoder is one of many methods in use for unsupervised feature learning~\cite{single-layer-networks-unsupervised-feature-learning}. It takes the form of a feedforward neural network, which consists of an input layer of N nodes, some number of hidden layers, and output layer of M nodes. The value of each layer $\vec{l_i}$ is determined with a weight matrix $W_i$ and a bias vector $b_i$: \[L_i = f(W_i l_{i-1} + b_i),\] where $f(\cdot)$ is a non-linearity like the sigmoid $f(x) = \frac{1}{1 + e^{-z}}$ applied component-wise.

An auto-encoder is such a neural network trained so as to minimize the error between the output layer and the input layer. If we use a short

\subsection{Manually-generated features}
\begin{table*}
\begin{tabular}{lp{0.6\textwidth}}
\hline
\textbf{Category} & \textbf{Description} \\ \hline
Position features & Piece location, potential moves per piece, material count, square control. \\
Move features & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move.
\hline
\end{tabular} 
\caption{List of chess-specific manual features per position.}
\end{table*}

As a first step towards classification, we selected chess-specific manual features from a given position-move pair.  The position-based features involve information such as where the pieces are on the board and potential moves, while the move-based features involve binary information such as if the move is a capturing move or a check.  We list all our hand-selected manual features in Table REF.

@@

TALK ABOUT THE AUTOENCODER EXPERIMENTS

@@

We then extended the tagging to also incorporate information about the player’s color because there is a bias. For the predictions from the SVM model, this modification improved recall at a minor cost to precision, resulting in a higher F1 score. In contrast, the random predictions got much worse: Both the precision and recall scores fell as a result of this change, which adds further bias and reduces the amount of data per tag (since games are now spread over twice as many tags).

TALK ABOUT AUTOENCODER RESULTS

TALK ABOUT USING AUTOENCODER inputs for the SVM

@@

\subsection{Classification Quality}
Future work should emphasize the quality of classifications. We demonstrated that general approaches may not be effective enough to be useful for complex games like chess. Therefore, future projects should seek improvements by better taking advantage of the chess-specific tools and libraries that are available like game engines and weight functions. If the classifier incorporated chess-specific features like those generated by chess engines, it could be much better at identifying the types of strategies that are available to the player. For example, a deep traversal of many future paths by a chess engine may reveal that an attack is possible (attackPossible) or that the player can force a pawn promotion (pawnPromotion). While a purely computational approach to identifying opportunities and strategies may provide higher-quality or more reliable assessments, a statistical approach may be more robust. %talk about what i mean by robust

One potential improvement that does not involve increased reliance on chess-specific technologies is to implement hierarchical tags so that matingAttack, doubleAttack, and similar tags imply a more general tag like attackPossible. By doing so, there will be a greater balance of positives and negatives, and labels that applied to games but were previously considered wrong are now considered correct. By incorporating hierarchical structure to tags, the system can jointly train and classify games.

Another possible improvement is to use a different classifier than SVM, such as the perceptron or maximum entropy classifiers. One issue with the SVM is that often the tags are not so clear-cut: What does it mean for example that an endgame is “good”? Rather, there are degrees of how good an endgame is to the player. Therefore, a softer approach like maximum entropy may produce better results.

\subsection{Generalization to Other Games.}
One goal of this project was to develop a statistical machine learning approach that was general enough to be easily adapted to other applications like StarCraft or Checkers. A future work can explore this possibility by simply creating features for other applications.
